---
title: "Lecture 2A - Monte Carlo - Work in Process"
author: "Robert Settlage"
authorbox: false
slug: "Lecture 2A"
date: 2020-11-05
publishdate: 2020-11-05
draft: false
categories:
- Lecture2
tags:
- Monte Carlo
output:
  html_document: default
  fig_caption: yes
  fig_height: 6
  fig_width: 7
  pdf_document: null
  classoption: landscape
header-includes: \setlength\parindent{24pt} \usepackage{MnSymbol} \usepackage{mathrsfs}
---

## Monte Carlo coming soon ... 

<!--more-->

```{r}
plot(cars) 
```

### Example problem - Darts for pi(e)

Monte Carlo methods are ones that involve repeated random sampling to estimate unknown parameters.  As an example, we can use repeated random sampling from uniform distributions to estimate pi.  Consider playing darts where we have inscribed a circle within the area given by a dart board.  If the circle and square are centered at c(0,0) where the circle has radius r=1 while the square has sides of length d=2, we can calculate the probablity of hitting the circle vs the square as the ratio of areas given by:

\begin{eqnarray}
\frac{C}{S} &=& \frac{\pi \ast r^2}{(2 \ast r)^2} \\
    &=& \frac{\pi}{4}
\end{eqnarray}

For starters, we will do this using loops and not worry about optimizing via apply functions, memory management or other tricks.

```{r}
options(digits=15)
cat("pi=",pi,"\n",sep="")
set.seed(123456)
while()
```

